% \input{\pSections "sec-writ-large"}

\section{Writ Large}

%     %     %     %     %     %     %     %     %
\subsection{Radar Images of Satellite}
\begin{frame}\frametitle{Sealing the Mesh}
Boo
\end{frame}
		%


%     %     %     %     %     %     %     %     %
\subsection{Approximation}
\begin{frame}{Fourier and Extensions to 2- and 3-D}
    \centering
    $\fourier$
    \vspace{1em}
    \textit{From periodic functions in 1D to radial and angular decompositions in 2D and 3D.}
\end{frame}
	
\begin{frame}{Why We Love Fourier for Smooth Functions}
    \textbf{Smooth Functions, Beautiful Representations}
    \begin{itemize}
        \item \textbf{Weierstrass Approximation Theorem:} 
        Any continuous function on \([a, b]\) can be uniformly approximated by polynomials. Fourier provides a similar approximation, using trigonometric bases instead of polynomials.
        
        \item \textbf{Riesz-Fischer Theorem:} 
        Fourier coefficients \((a_n, b_n)\) belong to \(l^2\) space, guaranteeing convergence in the \(L^2\) sense. This bridges the gap between smoothness and square-integrability.

        \item \textbf{Uniform Convergence for Smooth Periodic Functions:} 
        For sufficiently smooth functions (\(C^\infty\) or \(C^k\)), Fourier series converge uniformly, ensuring no oscillatory artifacts (Gibbs phenomenon disappears).

        \item \textbf{Orthogonality of Basis:} 
        Sines and cosines form an orthogonal basis in \(L^2\), enabling direct computation of coefficients via projection (Parseval's theorem quantifies energy distribution).

        \item \textbf{Spectral Insights:} 
        Decomposes a function into frequency components, making smoothness and structure explicit. High smoothness = rapid Fourier coefficient decay.

        \item \textbf{Compact Representation:} 
        Smooth functions require fewer terms for accurate approximation. High efficiency for practical computation and storage.

        \item \textbf{Universality:} 
        Fourier's reach extends to PDEs, signal processing, and quantum mechanics. Smooth functions unlock the full power of these tools.
    \end{itemize}
    \vspace{1em}
    \textbf{Takeaway:} Fourier connects smoothness, convergence, and representation, offering unmatched clarity and utility for periodic and localized phenomena.
\end{frame}

\begin{frame}{Fourier and Extensions to 2- and 3D}
    \centering
    \renewcommand{\arraystretch}{2.0}
    \setlength{\tabcolsep}{10pt} % Adjust column spacing
    
    \begin{tabular}{@{} l c @{}} % Align left and center without extra spacing
        \textcolor{blue}{\textbf{1D:}} & 
        \( \displaystyle f(\theta) = \sum_{n=-\infty}^{\infty} a_m \fourier \) \\[1em]

        \textcolor{blue}{\textbf{2D:}} & 
        \( \displaystyle f(r, \theta) = \sum_{n=1}^{\infty}\sum_{m=-n}^{n} R_n^m(r) fourier  \) \\[1em]

        \textcolor{blue}{\textbf{3D:}} & 
        \( \displaystyle 
        f(r, \theta, \phi) = 
        \sum_{n=1}^{\infty}\sum_{m=-n}^{n} 
        \sqrt{\frac{(2l+1)(l-m)!}{4\pi (l+m)!}} 
        P_l^m(\cos \theta) \fourier 
        \)
    \end{tabular}
\end{frame}
	
\begin{frame}{Fourier and Extensions to 2- and 3D}
    \begin{itemize}
        \item \href{https://mathworld.wolfram.com/FourierSeries.html}{\textbf{1D Fourier Series:}} Decomposes a periodic function \(f(\theta)\) into a sum of complex exponentials with coefficients \(a_n\) capturing the amplitudes of each frequency component.
        \item \href{https://mathworld.wolfram.com/ZernikePolynomial.html}{\textbf{2D Fourier-Bessel:}} Extends Fourier analysis to two dimensions using radial functions \(R_n^m(r)\), often employed in circular domains or optical applications.
        \item \href{https://mathworld.wolfram.com/SphericalHarmonic.html}{\textbf{3D Spherical Harmonics:}} Represents functions on a sphere using harmonics \(Y_l^m(\theta, \phi)\) and radial components \(R_l(r)\), crucial in fields like quantum mechanics and gravitational modeling.
    \end{itemize}
\end{frame}


%     %     %     %     %     %     %     %     %
\subsection{Quality of Fit}

\endinput  %  ==  ==  ==  ==  ==  ==  ==  ==  ==
