% \input{\pSections/sec-first.tex}

\section{Introduction}

The Radial Alignment and Vectorized Thrust Orientation in Time (RaVThOughT) navigation reference frame proposes a novel approach to spacecraft guidance and control by simplifying local maneuvering while maintaining precision. This document analyzes its contributions, compares it with traditional methods, explores its mathematical foundation, and suggests related research topics.

\section{Key Contributions of RaVThOughT}

\subsection{Simplification of Guidance Algorithms}
- Decouples local maneuvering from gravitational effects.
- Employs quadratic interpolation for gravitational transformations over 100-second windows.
- Reduces computational load by an order of magnitude compared to traditional methods.

\subsection{Machine Learning Compatibility}
- Simplifies state representation for better action-outcome relationships.
- Accelerates training of machine learning models while preserving physical meaning.
- Facilitates reinforcement learning in spacecraft guidance applications.

\subsection{Enhanced Multi-Vehicle Coordination}
- Introduces hierarchical compound frames for managing constellations and formation flying.
- Simplifies relative motion and collision avoidance tasks.
- Scales efficiently for cooperative missions.

\subsection{Built-in Error Detection}
- Employs a left-handed coordinate system, unique among standard reference frames.
- Prevents subtle errors by clearly distinguishing from traditional right-handed systems.

\section{Comparison with Existing Reference Frames}

Traditional reference frames have distinct strengths and limitations:

\begin{table}[h!]
\centering
\begin{tabular}{|l|l|l|}
\hline
\textbf{Frame} & \textbf{Strengths} & \textbf{Limitations} \\
\hline
Earth-Centered Inertial (ECI) & Accurate for long-term orbital evolution & Poor intuition for local maneuvers \\
Earth-Centered Earth-Fixed (ECEF) & Ground-relative operations & Rotational complexity for orbital navigation \\
Local-Vertical-Local-Horizontal (LVLH) & Intuitive for relative motion & Computationally intensive for long-term predictions \\
Radial-Space-Walk (RSW) & Simplifies relative motion in orbital planes & Complex for machine learning and multi-vehicle coordination \\
\hline
\end{tabular}
\caption{Comparison of traditional reference frames}
\end{table}

RaVThOughT bridges these gaps by combining the simplicity of vectorized maneuvers with manageable gravitational models, making it ideal for discrete guidance algorithms and machine learning.

\section{Mathematical Framework}

\subsection{Local Reference Frame}
- Left-handed coordinate system anchored to spacecraft features:
  - \(+X\): Primary thrust vector.
  - \(+Z\): Antenna axis (inward).
  - \(+Y\): Completes the left-handed system \((\mathbf{X} \times \mathbf{Z})\).

\subsection{State Representation}
Each RaVThOughT point specifies:
\begin{itemize}
  \item \textbf{Position:} \(\mathbf{r} = (x, y, z)\).
  \item \textbf{Velocity:} \(\mathbf{v} = (v_x, v_y, v_z)\).
  \item \textbf{Orientation:} \(\boldsymbol{\theta} = (\theta_x, \theta_y, \theta_z)\).
  \item \textbf{Angular Rates:} \(\boldsymbol{\omega} = (\omega_x, \omega_y, \omega_z)\).
  \item \textbf{Time:} Absolute mission time, \(t_{\text{abs}}\).
\end{itemize}

\subsection{Gravitational Rectification}
Gravitational effects are interpolated using quadratic regression:
\[
\mathbf{r}(t) = \mathbf{r}_0 + \mathbf{v}_0 t + \frac{1}{2} \mathbf{a}_g t^2,
\]
where \(\mathbf{a}_g\) is the gravitational acceleration, approximated as constant over 100 seconds. For typical Low Earth Orbit (LEO) conditions:
\begin{itemize}
  \item Gravitational acceleration variation \(\Delta g \approx 0.01\%\).
  \item Errors remain within centimeter-scale precision.
\end{itemize}

\subsection{Progress Metrics}
Linear progress scaling simplifies tracking:
\[
\text{Progress} = \frac{t}{100}, \quad \text{for } t \in [0, 100] \text{ seconds}.
\]

\section{Topics for Literature Search}

\subsection{Reference Frames in Orbital Mechanics}
- Foundational works on ECI, ECEF, LVLH, and RSW frames.
- Advancements in relative motion dynamics (e.g., Hill-Clohessy-Wiltshire equations).

\subsection{Simplified Orbital Mechanics}
- Interpolation methods (e.g., quadratic, cubic) for gravitational effects.
- Time-stepping techniques for short-term navigation.

\subsection{Machine Learning in Spacecraft Guidance}
- Applications of machine learning in guidance, navigation, and control (GNC).
- Reinforcement learning for orbital maneuver planning.

\subsection{Formation Flying and Constellation Management}
- Algorithms for multi-vehicle coordination and collision avoidance.
- Synchronization techniques for large constellations.

\subsection{Error Detection in Guidance Systems}
- Impact of coordinate system errors on simulations.
- Novel approaches to error prevention through frame design.


\endinput  %  ==  ==  ==  ==  ==  ==  ==  ==  ==
